---
title: "p8105_hw2_mk4208"
output: github_document
author: Meeraj Kothari 
date: 3/10/2019
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##### Loading required packages 

```{r}
library(tidyverse)
library(readxl)
```

## Problem 1 
Importing Mr. Trash Wheel sheet and omitting non-data entries. 
Rows that do not include dumpster-specific data have been omitted. 
Number of sports balls have been rounded the nearest integer and the result has been converted to a numeric variable. 

```{r}
trashwheel =
  read_excel(path="./Data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
             sheet = 1,
             range ="A2:N408") %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(sports_balls = round(sports_balls, digits = 0)) %>% 
  mutate(sports_balls = as.integer(sports_balls))
```

# Precipitation data for 2018:
Rows without precipitation data have been omitted. 
Year variable has been added.

```{r}
precipitation_2018 = 
  read_excel(path="./Data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", 
                        sheet = 3,
                        range ="A2:B14") %>% 
  janitor::clean_names()  %>%
  drop_na(total) %>% 
  mutate(year = 2018)
```

# Precipitation data for 2017: 
Rows without precipitation data have been omitted. 
Year variable has been added. 

```{r}
precipitation_2017 = 
  read_excel(path="./Data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", 
                        sheet = 4,
                        range ="A2:B14") %>% 
  janitor::clean_names()  %>%
  drop_na(total) %>% 
  mutate(year = 2017)
```

Precipitation data for 2018 and 2017 have been merged and month name has been converted to a character variable. 

```{r}
precipitation_full = full_join(precipitation_2017, precipitation_2018) %>% 
  mutate(month = month.name[month])
```

# Here is a short description of the 2 datasets: 
###### For the precipitation dataset:
* There are **`r nrow(precipitation_full)` observations**.
* It contains the following variables: **`r colnames(precipitation_full)`**.
* The total precipitation in the year 2018 was **`r sum(pull(precipitation_2018, total))`**. 

###### For the Mr. Trash Wheel dataset, 
* There are **`r nrow(trashwheel)` observations**.
* It contains the following variables: **`r colnames(trashwheel)`**.
* The median number of sports balls in a dumpster in 2017 was **`r median(pull(filter(trashwheel, year == "2017"), sports_balls))`**.

```{r}
pols = read_csv("./Data/fivethirtyeight_datasets/pols-month.csv") %>% 
  separate(mon,
    c("year", "month", "day"), 
    convert = TRUE) %>% 
  mutate(prez_dem = recode(prez_dem, '1' = "dem")) %>% 
  mutate(prez_gop = recode(prez_gop, '1' = "gop")) %>% 
  pivot_longer(
    c("prez_dem", "prez_gop"), 
    values_to = "president", 
    values_drop_na = TRUE
  ) %>% 
  mutate(month = month.name[month]) %>% 
  select(-day)

snp = read_csv("./Data/fivethirtyeight_datasets/snp.csv") %>% 
  separate(date, 
           c("month", "day", "year"),
           convert = TRUE) %>%
  mutate(month = month.name[month]) %>% 
  select(year, month, -day, everything())

unemployment = read_csv("./Data/fivethirtyeight_datasets/unemployment.csv") %>% 
  pivot_longer(Jan:Dec, 
    names_to = "month",
    values_to = "rate"
  ) %>% 
  mutate(month = match(month, month.abb)) %>% 
  mutate(month = month.name[month]) %>%
  mutate(Year = as.integer(Year)) %>%
  janitor::clean_names()

merged_data = pols %>% 
  left_join(snp, 
    by = c("year", "month") 
  ) %>% 
  left_join(
    unemployment,
    by = c("year", "month")
    )
```

# Problem 3: 
Loading the popular baby names dataset. Recoding the inconsistent ethinicity observations. Fixing the case structure of string varaibles. Removing duplicates using the distinct() function. 

```{r}
popular_baby_names = read_csv("./Data/Popular_Baby_Names.csv") %>%
  janitor::clean_names() %>% 
  mutate(ethnicity = recode(ethnicity, "ASIAN AND PACI" = "ASIAN AND PACIFIC ISLANDER",
                            "WHITE NON HISP" = "WHITE NON HISPANIC",
                            "BLACK NON HISP" = "BLACK NON HISPANIC")) %>% 
  mutate(childs_first_name = str_to_lower(childs_first_name),
         ethnicity = str_to_lower(ethnicity), 
         gender = str_to_lower(gender)) %>% 
  distinct()
```

Following table shows the rank in popularity of the name "Olivia" as a female baby name over time. Rows contain ethnicities and columns contain year. 

```{r} 
popular_baby_names %>% 
  filter(childs_first_name == "olivia") %>% 
  select(-count) %>%
  pivot_wider(
    names_from = "year_of_birth",
    values_from = "rank"
  ) %>% 
  knitr::kable()
```

Follwing table shows the most popular name among male children over time.

```{r}
popular_baby_names %>% 
  filter(gender == "male") %>% 
  filter(rank == 1) %>%
  select(-count) %>% 
  pivot_wider(
    names_from = "year_of_birth",
    values_from = "childs_first_name"
  ) %>% 
  knitr::kable()
```

##### For male, white non-hipsanic children born in 2016: 
Following is a scatter plot showing the number of children with a name (y axis) against the rank in popularity of that name (x axis) 

```{r}
popular_baby_names %>% 
  filter(gender ==  "male", 
         ethnicity == "white non hispanic", 
         year_of_birth == 2016) %>%
  ggplot(aes(x = rank, y = count)) +
  geom_point(colour = "blue") + 
  geom_text(aes(label = childs_first_name),
            hjust = 1, 
            vjust = 1, 
            check_overlap = TRUE,
            size = 2.4)
```



