---
title: "p8105_hw2_mk4208"
output: github_document
author: Meeraj Kothari 
date: 3/10/2019
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##### Loading required packages 

```{r message=FALSE}
library(tidyverse)
library(readxl)
```

# Problem 1 
Importing Mr. Trash Wheel sheet and omitting non-data entries. 
Rows that do not include dumpster-specific data have been omitted. 
Number of sports balls have been rounded the nearest integer and the result has been converted to a numeric variable. 

```{r}
trashwheel =
  read_excel(path="./Data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
             sheet = 1,
             range ="A2:N408") %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(sports_balls = round(sports_balls, digits = 0)) %>% 
  mutate(sports_balls = as.integer(sports_balls))
```

## Here is a short description of the dataset: 
###### For the Mr. Trash Wheel dataset, 
* There are **`r nrow(trashwheel)` observations**.
* It contains the following variables: **`r colnames(trashwheel)`**.
* The median number of sports balls in a dumpster in 2017 was **`r median(pull(filter(trashwheel, year == "2017"), sports_balls))`**.

## Precipitation data for 2018:
Rows without precipitation data have been omitted. 
Year variable has been added.

```{r}
precipitation_2018 = 
  read_excel(path="./Data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", 
                        sheet = 3,
                        range ="A2:B14") %>% 
  janitor::clean_names()  %>%
  drop_na(total) %>% 
  mutate(year = 2018)
```

## Precipitation data for 2017: 
Rows without precipitation data have been omitted. 
Year variable has been added. 

```{r}
precipitation_2017 = 
  read_excel(path="./Data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", 
                        sheet = 4,
                        range ="A2:B14") %>% 
  janitor::clean_names()  %>%
  drop_na(total) %>% 
  mutate(year = 2017)
```

Precipitation data for 2018 and 2017 have been merged and month name has been converted to a character variable. 

```{r message=FALSE}
precipitation_full = full_join(precipitation_2017, precipitation_2018) %>% 
  mutate(month = month.name[month])
```

## Here is a short description of the dataset: 
###### For the precipitation dataset:
* There are **`r nrow(precipitation_full)` observations**.
* It contains the following variables: **`r colnames(precipitation_full)`**.
* The total precipitation in the year 2018 was **`r sum(pull(precipitation_2018, total))`**. 

---------------------------------------------------------------------------------------------

## Loading `pols`, `snp` and `unemployment` datasets...

### For the `pols` dataset:
The month variable is separated into "year", "month" and "day". A new variable "President" has been added taking values from prez_dem and prez_gop. The values of the month variable have been converted to characters. Day variable has been removed. 


```{r message=FALSE}
pols = read_csv("./Data/fivethirtyeight_datasets/pols-month.csv") %>% 
  separate(mon,
    c("year", "month", "day"), 
    convert = TRUE) %>% 
  mutate(prez_dem = recode(prez_dem, '1' = "dem")) %>% 
  mutate(prez_gop = recode(prez_gop, '1' = "gop", '2' = "gop")) %>% 
  pivot_longer(
    cols = c("prez_dem", "prez_gop"), 
    values_to = "president", 
    values_drop_na = TRUE
  ) %>% 
  mutate(month = month.name[month]) %>% 
  select(-day, -name)
```

#####  Here is a short description of the dataset: 
* There are **`r nrow(pols)` observations** and **`r ncol(pols)` columns**.
* The range of years is from **`r min(pull(pols, year))` to `r max(pull(pols, year))`**.
* It contains the following variables: **`r colnames(pols)`**.


### For the `snp` dataset:
The date variable is separated into "month", "day" and "year". The values of the month variable have been converted to characters. The year and month variables have been rearranged. Day variable has been removed. 

```{r message=FALSE}
snp = read_csv("./Data/fivethirtyeight_datasets/snp.csv") %>% 
  separate(date, 
           c("month", "day", "year"),
           convert = TRUE) %>%
  mutate(month = month.name[month]) %>% 
  select(year, month, -day, everything())
```

##### Here is a short description of the dataset: 
* There are **`r nrow(snp)` observations** and **`r ncol(snp)` columns**.
* The range of years is from **`r min(pull(snp, year))` to `r max(pull(snp, year))`**.
* It contains the following variables: **`r colnames(snp)`**.

### For the `unemployment` dataset:
A new variable month has been created taking values from the Jan to Dec column names. The month variable has been formatted to character variable using month,name. Year variable has been converted to integer.  

```{r message=FALSE}
unemployment = read_csv("./Data/fivethirtyeight_datasets/unemployment.csv") %>% 
  pivot_longer(Jan:Dec, 
    names_to = "month",
    values_to = "rate"
  ) %>% 
  mutate(month = match(month, month.abb)) %>%
  mutate(month = month.name[month]) %>%
  mutate(Year = as.integer(Year)) %>%
  janitor::clean_names()
```

##### Here is a short description of the dataset: 
* There are **`r nrow(unemployment)` observations** and **`r ncol(unemployment)` columns**.
* The range of years is from **`r min(pull(unemployment, year))` to `r max(pull(unemployment, year))`**.
* It contains the following variables: **`r colnames(unemployment)`**.

----------------------------------------------------------------------------------------------

The following code chunk merges `snp` into `pols` and then `unexmployment` into the result. 

```{r}
merged_data = pols %>% 
  left_join(snp, 
            by = c("year", "month")) %>%
  left_join(unemployment, 
            by = c("year", "month"))
```

##### Here is a short description of the dataset: 
###### For the merged dataset:
* There are **`r nrow(merged_data)` observations** and **`r ncol(merged_data)` columns**.
* The range of years is from **`r min(pull(merged_data, year))` to `r max(pull(merged_data, year))`**.
* It contains the following variables: **`r colnames(merged_data)`**.

# Problem 3: 
Loading the popular baby names dataset. Recoding the inconsistent ethinicity observations. Fixing the case structure of string varaibles. Removing duplicates using the distinct() function. 

```{r message=FALSE}
popular_baby_names = read_csv("./Data/Popular_Baby_Names.csv") %>%
  janitor::clean_names() %>% 
  mutate(ethnicity = recode(ethnicity, "ASIAN AND PACI" = "ASIAN AND PACIFIC ISLANDER",
                            "WHITE NON HISP" = "WHITE NON HISPANIC",
                            "BLACK NON HISP" = "BLACK NON HISPANIC")) %>% 
  mutate(childs_first_name = str_to_lower(childs_first_name),
         ethnicity = str_to_lower(ethnicity), 
         gender = str_to_lower(gender)) %>% 
  distinct()
```

Following table shows the rank in popularity of the name "Olivia" as a female baby name over time. Rows contain ethnicities and columns contain year. 

```{r} 
popular_baby_names %>% 
  filter(childs_first_name == "olivia") %>% 
  select(-count) %>%
  pivot_wider(
    names_from = "year_of_birth",
    values_from = "rank"
  ) %>% 
  knitr::kable()
```

Follwing table shows the most popular name among male children over time.

```{r}
popular_baby_names %>% 
  filter(gender == "male") %>% 
  filter(rank == 1) %>%
  select(-count) %>% 
  pivot_wider(
    names_from = "year_of_birth",
    values_from = "childs_first_name"
  ) %>% 
  knitr::kable()
```

##### For male, white non-hispanic children born in 2016: 
Following is a scatter plot showing the number of children with a name (y axis) against the rank in popularity of that name (x axis) 

```{r}
popular_baby_names %>% 
  filter(gender ==  "male", 
         ethnicity == "white non hispanic", 
         year_of_birth == 2016) %>%
  ggplot(aes(x = rank, y = count)) +
  geom_point(colour = "blue") + 
  geom_text(aes(label = childs_first_name),
            hjust = 1, 
            vjust = 1, 
            check_overlap = TRUE,
            size = 2.4)
```



